---
title: "SysReview_phytoplankton-water-level-change"
author: "Kathryn K. Hoffman"
format: html
editor: visual
---

## Quarto

This markdown:

1.  reads in the extracted data from the systematic literature review.
2.  applies 1_qaqc_data_sysreview.R to exclude any sites that were erroneously extracted, to de-duplicate any sites/data that were included in this review in multiple studies, and to fix assorted typos and mistakes in the raw extracted data.
3.  applies 2_extract_drivers.R to create new columns for phytoplankton and cyanobacterial drivers, responses, and measured variables.
4.  applies 3_proportion_WL.R to calculate proportional change in water level for the logistic regression.
5.  creates all of the manuscript figures.
6.  creates all of the supplemental information (SI) figures.

```{r}
  library(dplyr)
  library(tidyr)
  library(janitor)
  library(readr)
  library(here)
  
```

## Libraries

Install and load dependencies.

```{r packages}
#| message: false
#| warning: false

if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")

pacman::p_load(janitor, tidyverse, ggplot2, dplyr, tidyr, naniar, plotly, patchwork, here, #plotting and data manipulation packages
               rnaturalearth, rnaturalearthdata, mapview, sf, #mapping packages
               rstatix, rfriend) #chi-square analysis packages

#MASS needs to be installed, but is installed and loaded later because of conflicts with dplyr
```

```{r read in and qaqc data}
#read in and qaqc data
source(here::here("Scripts/1_qaqc_data_sysreview.R"))

results <- qaqc_data_sysreview("Data/sys_review_extracted_data.csv")

data <- results$data
excluded_papers <- results$excluded_papers

write.csv(data, file = here("Data/qaqc_sysreviewdata.csv"))

```

## Dataset summary statistics

```{r quick study info}

#number of waterbodies

reservoirs <- data |> select(c(res_name)) |> unique()
countries <- data |> select(c(country)) |> unique()
studies <- data |> select(c(name_yr)) |> unique()

summary_counts <- tibble(
  metric = c("reservoirs", "countries", "studies"),
  value = c(
    dplyr::n_distinct(data$res_name),
    dplyr::n_distinct(data$country),
    dplyr::n_distinct(data$name_yr)
  )
)
summary_counts
#there are a total of 3 TGRs, so there are 64-2 reservoirs, = 62 total lakes!
#25 countries - combined Argentina/Uruguay, Estonia/Russia, one blank = 22 total countries
#60 studies, some duplicate authors but no duplicate studies.



both_measured <- data |>
  select(res_name, increase_phyto_response, increase_cyano_response,
         decrease_phyto_response, decrease_cyano_response) |>
  mutate(across(
    c(decrease_phyto_response, increase_phyto_response,
      decrease_cyano_response, increase_cyano_response),
    ~ .x |>
      trimws() |>
      na_if("") |>
      na_if("Not measured")
  )) |>
  mutate(test = ifelse(
    ((!is.na(decrease_phyto_response) | !is.na(increase_phyto_response)) &
     (!is.na(decrease_cyano_response) | !is.na(increase_cyano_response))),
    "Both", "one"
  )) |>
  filter(test == "Both")

result <- data %>%
  # clean missing values
  mutate(across(
    c(decrease_phyto_response, decrease_cyano_response, phys_drive),
    ~ .x %>%
      trimws() %>%
      na_if("") %>%
      na_if("Not measured")
  )) %>%
  filter(
  grepl("decrease", decrease_phyto_response, ignore.case = TRUE) |
  grepl("decrease", decrease_cyano_response, ignore.case = TRUE)
) |> 
  
  # flag whether drivers mention turbidity / TSS / low light
  mutate(driver_flag = str_detect(
    tolower(phys_drive),
    "turbidity|total suspended solids|light"
  )) %>%
  select(decrease_phyto_response, increase_phyto_response, decrease_cyano_response, res_name, driver_flag, name_yr, phys_drive)
                           
```

# Manuscript Figures

```{r extract drivers responses and measured vars}
source(here::here("Scripts/2_extract_drivers.R"))

data_wdrivers <- extract_drivers(data)
```

```{r create a folder for mansucript figures}
   manuscript_figures <- here("Manuscript_figures/")
if(!file.exists(manuscript_figures)){
    #make the folder
    dir.create(file.path(manuscript_figures))
  }
```

## Phyto and cyano drivers

```{r phyto drivers}

phyto_only <- data_wdrivers |> 
  filter(
    increase_phyto_response != "" & increase_phyto_response != "Not measured" |
    decrease_phyto_response != "" & decrease_phyto_response != "Not measured"
  )

chem_phyto_only_drivers <- phyto_only  |> 
  select(matches("^(cond|do|nh4|no2|no3|orp|ph|srp|tn|tp).*(?:_vars|_drive)$"), res_name, name_yr) |> 
  select(-c(res_name, name_yr, phys_vars, phys_drive, phyto_micros_vars, dom_vars, dom_drive)) #|> 
  #rename_with(~ paste0("phyto_", .))

chem_phyto_sums <- chem_phyto_only_drivers %>%
  summarise(across(everything(), sum, na.rm = TRUE))

chem_phyto_long <- chem_phyto_sums %>%
  pivot_longer(cols = everything(),
               names_to = "variable_type",
               values_to = "count") %>%
  separate(variable_type, into = c("variable", "type"), sep = "_(?=[^_]+$)")

chemicalvars_phytos <- chem_phyto_long %>%
  mutate(type = recode(type, vars = "meas")) %>%
  pivot_wider(
    names_from = type,
    values_from = count) |> 
  rename(id = drive, 
         names = variable) |> 
  mutate(notid = meas - id, 
         prop_id = id / meas)

chem <- chemicalvars_phytos |> 
  select(names, notid, id, prop_id) |> 
  arrange(desc(prop_id)) |> 
  mutate(names = factor(names, levels = names)) |> 
  select(-prop_id) |> 
  pivot_longer(!names, names_to = "driver", values_to = "count") |> 
  ggplot(aes(fill = driver, y = count, x = names)) +
  geom_bar(position = position_fill(reverse = TRUE), stat = "identity") +
  geom_text(
    aes(label = count),
    position = position_fill(vjust = 0.5, reverse = TRUE),
    color = "black",
    size = 3
  ) +
  scale_fill_manual(
    values = c("#99bbab", "#ddefcd"),
    labels = c("Important", "Not Important"),
    name = "Phytoplankton:\nDid authors identify\nthis driver as important?"
  ) +
  ggtitle("Chemical Variables") +
  theme_bw() +
  labs(y = "") +
  theme(
    axis.text.x = element_text(hjust = .5, size = 6),
    axis.title.x = element_blank(),
    axis.title.y = element_text(angle = 90, vjust = .5, size = 12)
  )

phys_phyto_only_drivers <- phyto_only  |> 
  select(matches("^(water|transparency|turb|light|tss_tds).*(?:_vars|_drive)$"), res_name, name_yr) |> select(-c(res_name, name_yr)) |> 
  #rename_with(~ paste0("phyto_", .))
mutate(total_wl = ifelse(water_level_drive == 1 & water_level_vars == 0, 1, water_level_vars), 
       water_level_vars = total_wl) |> 
  select(-c(total_wl))
  
phys_phyto_sums <-phys_phyto_only_drivers %>%
  summarise(across(everything(), sum, na.rm = TRUE))

phys_phyto_long <- phys_phyto_sums %>%
  pivot_longer(cols = everything(),
               names_to = "variable_type",
               values_to = "count") %>%
  separate(variable_type, into = c("variable", "type"), sep = "_(?=[^_]+$)")

physicalvars_phytos <- phys_phyto_long %>%
  mutate(type = recode(type, vars = "meas")) %>%
  pivot_wider(
    names_from = type,
    values_from = count) |> 
  rename(id = drive, 
         names = variable) |> 
  mutate(notid = meas - id, 
         prop_id = id / meas)

phys <- physicalvars_phytos |> 
  # select(names, measured, sum) |> 
  # filter(!names == "other_phys_vars") |> 
  # rename(id = sum) |> 
  # mutate(notid = measured - id,
  #        prop_id = id / measured) |> 
  select(names, notid, id, prop_id) |> 
  arrange(desc(prop_id)) |> 
  mutate(names = factor(names, levels = names)) |> 
  select(-prop_id) |> 
  pivot_longer(!names, names_to = "driver", values_to = "count") |> 
  ggplot(aes(fill = driver, y = count, x = names)) +
  geom_bar(position = position_fill(reverse = TRUE), stat = "identity") +
  geom_text(
    aes(label = count),
    position = position_fill(vjust = 0.5, reverse = TRUE),
    color = "black",
    size = 3
  ) +
  scale_fill_manual(
    values = c("#99bbab", "#ddefcd"),
    labels = c("Important", "Not Important"),
    name = "Phytoplankton:\nDid authors identify\nthis driver as important?"
  ) +
  ggtitle("Physical Variables") +
  theme_bw() +
  labs(y = "") +
  theme(
    axis.text.x = element_text(hjust = .5, size = 6),
    axis.title.x = element_blank(),
    axis.title.y = element_text(angle = 90, vjust = .5, size = 12)
  )

phytos <- phys/chem
phytos
```

```{r cyano drivers}
cyano_only <- data_wdrivers |> 
  filter(
    increase_cyano_response != "" & increase_cyano_response != "Not measured" |
    decrease_cyano_response != "" & decrease_cyano_response != "Not measured"
  )

chem_cyano_only_drivers <- cyano_only  |> 
  select(matches("^(cond|do|nh4|no2|no3|orp|ph|srp|tn|tp).*(?:_vars|_drive)$"), res_name, name_yr) |> 
  select(-c(res_name, name_yr, phys_vars, phys_drive, phyto_micros_vars, dom_vars, dom_drive)) #|> 
  #rename_with(~ paste0("cyano_", .))

chem_cyano_sums <- chem_cyano_only_drivers %>%
  summarise(across(everything(), sum, na.rm = TRUE))

chem_cyano_long <- chem_cyano_sums %>%
  pivot_longer(cols = everything(),
               names_to = "variable_type",
               values_to = "count") %>%
  separate(variable_type, into = c("variable", "type"), sep = "_(?=[^_]+$)")

chemicalvars_cyanos <- chem_cyano_long %>%
  mutate(type = recode(type, vars = "meas")) %>%
  pivot_wider(
    names_from = type,
    values_from = count) |> 
  rename(id = drive, 
         names = variable) |> 
  mutate(notid = meas - id, 
         prop_id = id / meas)

chemvars_cyanos_drivers <- chemicalvars_cyanos |> 
  select(names, notid, id, prop_id) |> 
  arrange(desc(prop_id)) |> 
  mutate(names = factor(names, levels = names)) |> 
  select(-prop_id) |> 
  pivot_longer(!names, names_to = "driver", values_to = "count") |> 
  ggplot(aes(fill = driver, y = count, x = names)) +
  geom_bar(position = position_fill(reverse = TRUE), stat = "identity") +
  geom_text(
    aes(label = count),
    position = position_fill(vjust = 0.5, reverse = TRUE),
    color = "black",
    size = 3
  ) +
  scale_fill_manual(
    values = c("#4dd0e1","#b2ebf2"),
    labels = c("Important", "Not Important"),
    name = "Cyanobacteria:\nDid authors identify\nthis driver as important?"
  ) +
  ggtitle("Chemical Variables") +
  theme_bw() +
  labs(y = "") +
  theme(
    axis.text.x = element_text(hjust = .5, size = 6),
    axis.title.x = element_blank(),
    axis.title.y = element_text(angle = 90, vjust = .5, size = 12)
  )

phys_cyano_only_drivers <- cyano_only  |> 
  select(matches("^(water|transparency|turb|light|tss_tds).*(?:_vars|_drive)$"), res_name, name_yr) |> select(-c(res_name, name_yr)) |> 
  #rename_with(~ paste0("cyano_", .))
mutate(total_wl = ifelse(water_level_drive == 1 & water_level_vars == 0, 1, water_level_vars), 
       water_level_vars = total_wl) |> 
  select(-c(total_wl))
  
phys_cyano_sums <-phys_cyano_only_drivers %>%
  summarise(across(everything(), sum, na.rm = TRUE))

phys_cyano_long <- phys_cyano_sums %>%
  pivot_longer(cols = everything(),
               names_to = "variable_type",
               values_to = "count") %>%
  separate(variable_type, into = c("variable", "type"), sep = "_(?=[^_]+$)")

physicalvars_cyanos <- phys_cyano_long %>%
  mutate(type = recode(type, vars = "meas")) %>%
  pivot_wider(
    names_from = type,
    values_from = count) |> 
  rename(id = drive, 
         names = variable) |> 
  mutate(notid = meas - id, 
         prop_id = id / meas)

physvars_cyanos_drivers <- physicalvars_cyanos |> 
  # select(names, measured, sum) |> 
  # filter(!names == "other_phys_vars") |> 
  # rename(id = sum) |> 
  # mutate(notid = measured - id,
  #        prop_id = id / measured) |> 
  select(names, notid, id, prop_id) |> 
  arrange(desc(prop_id)) |> 
  mutate(names = factor(names, levels = names)) |> 
  select(-prop_id) |> 
  pivot_longer(!names, names_to = "driver", values_to = "count") |> 
  ggplot(aes(fill = driver, y = count, x = names)) +
  geom_bar(position = position_fill(reverse = TRUE), stat = "identity") +
  geom_text(
    aes(label = count),
    position = position_fill(vjust = 0.5, reverse = TRUE),
    color = "black",
    size = 3
  ) +
  scale_fill_manual(
    values = c("#4dd0e1","#b2ebf2"),
    labels = c("Important", "Not Important"),
    name = "Cyanobacteria:\nDid authors identify\nthis driver as important?"
  ) +
  ggtitle("Physical Variables") +
  theme_bw() +
  labs(y = "") +
  theme(
    axis.text.x = element_text(hjust = .5, size = 6),
    axis.title.x = element_blank(),
    axis.title.y = element_text(angle = 90, vjust = .5, size = 12)
  )

cyanos <- physvars_cyanos_drivers/chemvars_cyanos_drivers
cyanos

combined <- (physvars_cyanos_drivers/chemvars_cyanos_drivers) | (phys / chem)
combined <- ((phys/physvars_cyanos_drivers) | (chem/chemvars_cyanos_drivers)) + plot_layout(guides = "collect") &
  theme(
    legend.position = "right",
    legend.box = "vertical"
  )
combined

ggsave(paste0(here(),"/Manuscript_figures/drivers_cyanophyto.jpg"), plot = combined, height = 4, width = 8, units = "in")
```

## Map

Make map figure.

```{r map figure}

na_strings <- c("NA", "N A", "N / A", "N/A", "N/ A", "Not Available", "NOt available", "", "not reported")

 latlong <- data %>%
  mutate(
    study_years = ifelse(
      grepl("36 \\(satellite data\\); 12 \\(field data\\)", study_years),"12", 
      ifelse(grepl("36 \\(satellite data\\); 2 \\(field data\\)", study_years),"2",study_years)), 
    study_years = as.numeric(study_years)) |> 
  mutate(depth_m = ifelse(!is.na(fullpond_maxdepth), fullpond_maxdepth, obs_maxdepth), 
         depth_m = as.numeric(depth_m), 
         years_studied = ifelse(study_years == 1, "1", 
                                ifelse(study_years >1 & study_years <=5, "2-5", 
                                       ifelse(study_years >5 & study_years <= 10, "6-10", 
                                              ifelse(study_years > 10, "more than 10", NA)))), 
         years_studied = factor(years_studied, levels = c("1", "2-5", "6-10", "more than 10"))) |> 
  select("lat_dd", "long_dd", "res_name", "decrease_phyto_response", "increase_phyto_response", "decrease_cyano_response", "increase_cyano_response", "depth_m", "trophic_status_reduced", "years_studied") |> 
  mutate(lat_dd = ifelse(lat_dd == 103.466667 & res_name == "Zipingpu Reservoir", 30.983333, lat_dd), 
         long_dd = ifelse(long_dd == 30.983333 & res_name == "Zipingpu Reservoir", 103.466667, long_dd)) |> 
  mutate(long_dd = ifelse(long_dd == 6.77888889 & res_name == "Serra Serrada", -6.77888889, long_dd)) |> 
   # mutate(lat_dd = ifelse(res_name == "Kuibyshev Reservoir", 53.462857, lat_dd), 
   #        long_dd = ifelse(res_name == "Kuibyshev Reservoir", 49.156064, long_dd)) |> 
    replace_with_na_all(condition = ~.x %in% na_strings) |> 
    subset(!is.na(long_dd)) |>  
  subset(!is.na(lat_dd)) |> 
  mutate(increase_phyto_response = ifelse(increase_phyto_response == "Not measured"|increase_phyto_response == ""|(grepl("Other", increase_phyto_response)), NA, increase_phyto_response)) |>
    mutate(decrease_phyto_response = ifelse(decrease_phyto_response == "Not measured"|decrease_phyto_response == ""|(grepl("Other", decrease_phyto_response)), NA, decrease_phyto_response)) |>
   mutate(increase_cyano_response = ifelse(increase_cyano_response == "Not measured"|increase_cyano_response == ""|(grepl("Other", increase_cyano_response)), NA, increase_cyano_response)) |>
   mutate(decrease_cyano_response = ifelse(decrease_cyano_response == "Not measured"|decrease_cyano_response == ""|(grepl("Other", decrease_cyano_response)), NA, decrease_cyano_response)) |>
  filter(!(is.na(decrease_phyto_response) & is.na(increase_phyto_response) & is.na(decrease_cyano_response) & is.na(increase_cyano_response))) %>%
  mutate(increase_decrease_mosaic = ifelse(((!is.na(decrease_phyto_response) | !is.na(decrease_cyano_response)) & (!is.na(increase_phyto_response) | !is.na(increase_cyano_response))),"Both",
                                           ifelse((!is.na(decrease_phyto_response) | !is.na(decrease_cyano_response)),"Decrease",
                                                  ifelse((!is.na(increase_phyto_response) | !is.na(increase_cyano_response)),"Increase", NA))))

world <- ne_countries(scale = "medium", returnclass = "sf")
#world <- map_data("world")

litdata_map <- latlong %>% 
  mutate(decrease_phyto_response = ifelse(decrease_phyto_response == "Not measured", NA, decrease_phyto_response), 
         increase_phyto_response = ifelse(increase_phyto_response == "Not measured", NA, increase_phyto_response), 
         decrease_cyano_response = ifelse(decrease_cyano_response == "Not measured", NA, decrease_cyano_response), 
         increase_cyano_response = ifelse(increase_cyano_response == "Not measured", NA, increase_cyano_response)) |> 
   mutate(mapvals = ifelse(((!is.na(decrease_phyto_response)|!is.na(increase_phyto_response)) & (!is.na(decrease_cyano_response)|!is.na(increase_cyano_response))), "Both",
                           ifelse(!is.na(decrease_phyto_response)|!is.na(increase_phyto_response), "Phytoplankton", 
                                 ifelse(!is.na(decrease_cyano_response)|!is.na(increase_cyano_response), "Cyanobacteria", NA)))) 

litdata_map_sf <- st_as_sf(litdata_map,
                           coords = c("long_dd", "lat_dd"),
                           crs = 4326,
                           remove = FALSE)

sitemap_trophic <- ggplot(world) +
  geom_sf(data = world, fill = "white") +
  #coord_sf(crs = 4269) +
  #scale_fill_manual(aesthetics = "white") +
  #geom_sf_label(stat = "sf_coordinates")+
  # geom_sf(data = litdata_map_sf,
  #         aes(shape = as.factor(mapvals),
  #             color = as.factor(increase_decrease_mosaic))) +
    geom_sf(data = litdata_map_sf, size = 2, alpha = 0.7,
          aes(color = as.factor(increase_decrease_mosaic))) +
  #scale_fill_manual(values = c("white", "#88CCEE"), name = "Countries") +
  scale_color_manual(values = c("#7A5FA8", "#C65A6A", "#3F7FA6"),
                     name = "Does water level \nincrease or decrease? \n") +
  scale_shape_manual(values = c(16, 17, 18),
                     name = "Variable")  +
  theme_minimal() + 
  guides(colour = guide_legend(override.aes = list(size=6))) +
  theme(legend.text  = element_text(size = 12),
        legend.title = element_text(size = 12))
sitemap_trophic

ggsave(paste0(here(),"/Manuscript_figures/sitemapposter.jpg"), plot = sitemap_trophic, height = 4, width = 10, units = "in")
```

The next chunk sources the function used to derive proportional water-level metrics.

```{r max depth and proportion water level}
#uses propWL_logreg function found in proportion_WL.R
source(here::here("Scripts/3_proportion_WL.R"))
boxplotdata2 <- propWL_logreg(data)
```

## Set up logistic regression for effect of water level change on phytoplankton and cyanobacteria communities

We ran the logistic regression three ways. In the manuscript, the logistic regression includes both water level increases and decreases. We also prioritized only water level increases (any lake that had both a water level increase and decrease was kept as only an increase) and subsequently prioritized only water level decreases (any lake that had both a water level increase and decrease was kept as only a decrease)

```{r logistic regression}
phytoincrease_logreg <- boxplotdata2 |> 
  dplyr::select(prop_alldelta_wl_decrease, prop_1delta_wl_increase, phyto_binary_increase, phyto_binary_decrease, phyto_response_wl_up, phyto_response_wl_down, name_yr, trophic_status_reduced, prop_alldelta_wl_increase, res_name) |> 
  mutate(phyto_binary_decrease = NA, 
         prop_alldelta_wl_decrease = NA, 
         prop_deltawl = prop_alldelta_wl_increase,
         response_binary = phyto_binary_increase,
         deltawl_type = "increase") |> 
  filter(!is.na(phyto_response_wl_up)) |> 
  filter(!is.na(prop_alldelta_wl_increase)) |> 
  select(deltawl_type, prop_alldelta_wl_increase, prop_alldelta_wl_decrease, phyto_binary_decrease, phyto_binary_increase, name_yr, trophic_status_reduced, prop_deltawl, response_binary, res_name)

phytodecrease_logreg <- boxplotdata2 |> 
  dplyr::select(prop_alldelta_wl_decrease, prop_1delta_wl_increase, phyto_binary_increase, phyto_binary_decrease, phyto_response_wl_up, phyto_response_wl_down, name_yr, trophic_status_reduced, prop_alldelta_wl_increase, res_name) |> 
  mutate(phyto_binary_increase = NA, 
         prop_alldelta_wl_increase = NA,
         prop_deltawl = prop_alldelta_wl_decrease,
         response_binary = phyto_binary_decrease,
         deltawl_type = "decrease") |> 
  filter(!is.na(phyto_response_wl_down)) |> 
  filter(!is.na(prop_alldelta_wl_decrease)) |> 
  select(deltawl_type, prop_alldelta_wl_increase, prop_alldelta_wl_decrease, phyto_binary_decrease, phyto_binary_increase, name_yr, trophic_status_reduced, prop_deltawl, response_binary, res_name)

phyto_logreg <- rbind(phytodecrease_logreg, phytoincrease_logreg)

phyto_logreg <- phyto_logreg |> 
  mutate(type = "phyto") |> 
  select(name_yr, res_name, trophic_status_reduced, prop_deltawl, response_binary, type)

cyanoincrease_logreg <- boxplotdata2 |> 
  dplyr::select(prop_alldelta_wl_decrease, cyano_binary_increase, cyano_binary_decrease, cyano_response_wl_up, cyano_response_wl_down, name_yr, trophic_status_reduced, prop_alldelta_wl_increase, res_name) |> 
  mutate(cyano_binary_decrease = NA, 
         prop_alldelta_wl_decrease = NA, 
         prop_deltawl = prop_alldelta_wl_increase,
         response_binary = cyano_binary_increase,
         deltawl_type = "increase") |> 
  filter(!is.na(cyano_response_wl_up)) |> 
  filter(!is.na(prop_alldelta_wl_increase)) |> 
  select(deltawl_type, prop_alldelta_wl_increase, prop_alldelta_wl_decrease, cyano_binary_decrease, cyano_binary_increase, name_yr, trophic_status_reduced, prop_deltawl, response_binary, res_name)

cyanodecrease_logreg <- boxplotdata2 |> 
  dplyr::select(prop_alldelta_wl_decrease, cyano_binary_increase, cyano_binary_decrease, cyano_response_wl_up, cyano_response_wl_down, name_yr, trophic_status_reduced, prop_alldelta_wl_increase, res_name) |> 
  mutate(cyano_binary_increase = NA, 
         prop_alldelta_wl_increase = NA,
         prop_deltawl = prop_alldelta_wl_decrease,
         response_binary = cyano_binary_decrease,
         deltawl_type = "decrease") |> 
  filter(!is.na(cyano_response_wl_down)) |> 
  filter(!is.na(prop_alldelta_wl_decrease)) |> 
  select(deltawl_type, prop_alldelta_wl_increase, prop_alldelta_wl_decrease, cyano_binary_decrease, cyano_binary_increase, name_yr, trophic_status_reduced, prop_deltawl, response_binary, res_name)

cyano_logreg <- rbind(cyanodecrease_logreg, cyanoincrease_logreg)

cyano_logreg <- cyano_logreg |> 
  mutate(type = "cyano") |> 
  select(name_yr, res_name, trophic_status_reduced, prop_deltawl, response_binary, type)

all_logreg <- rbind(phyto_logreg, cyano_logreg)


ggplot(all_logreg, aes(x = prop_deltawl, y = response_binary)) +
  geom_point(aes(color = trophic_status_reduced), alpha = 0.6, size = 4) +
  geom_smooth(
    method = "glm",
    method.args = list(family = "binomial"),
    se = FALSE,
    color = "black"
  ) +
  facet_wrap(~type) +
  theme_bw() +
  labs(
    x = "Proportional Water Level Change",
    y = "Probability of Increase"
  )

logregform_phyto <-  glm(formula=response_binary ~ prop_deltawl, data=phyto_logreg, family=binomial)
summary(logregform_phyto)

logregform_cyano <-  glm(formula=response_binary ~ prop_deltawl, data=cyano_logreg, family=binomial)
summary(logregform_cyano)

library(broom)

extract_summary <- function(model, data, model_name) {
  tidied <- broom::tidy(model)
  pval <- tidied$p.value[tidied$term == "prop_deltawl"]
  data_size <- nrow(data)
  
  tibble(
    model = model_name,
    n = data_size,
    pval = pval
  )
}

summary_tbl <- dplyr::bind_rows(
  extract_summary(logregform_phyto, phyto_logreg, "logregform_phyto"),
  extract_summary(logregform_cyano, cyano_logreg, "logregform_cyano")
)

summary_tbl

```

## Plot logistic regression with combined water level increase/decrease

This chunk will generate logistic regression plots for both phytoplankton and cyanobacteria.

```{r logreg plot}

all <- all_logreg %>% ggplot(aes(x = prop_deltawl, y = response_binary, color = as.factor(type), shape = as.factor(type))) +
  scale_color_manual(values = c("#4dd0e1", "#99bbab"), name = "")+
  scale_shape_manual(values = c(8, 20)) +
  geom_point(size = 3) +
  geom_smooth(method = "glm",
              method.args = list(family = "binomial"),
              se = FALSE, linewidth = 3) +
  #geom_jitter()+
 ylab("Response to changes in water level") +
  xlab("Proportion water level change") +
  ggtitle("Combined increases and decreases") + 
  theme_bw() +
  theme(axis.text.x = element_text(hjust = 1, size = 10),
        axis.title.y=element_text(angle=90, vjust = .5, size = 12),
        legend.title=element_blank())
all

ggsave(paste0(here(),"/Manuscript_figures/logreg.jpg"), plot = all, height = 4, width = 8, units = "in")

```

## Set up and run Chi-squared tests

These Chi-square tests assess the effect of increases and decreases in water level on phytoplankton and cyanobacteria (do they experience an increase, a decrease, or no change?).

```{r chisq tests of independence}
# Create contingency table
tab_cyano_all <- boxplotdata2 %>%
  pivot_longer(
    c(cyano_response_wl_up, cyano_response_wl_down),
    names_to = "water_direction_final",
    values_to = "response_final"
  ) %>%
  filter(!is.na(response_final)) %>%
  mutate(response_final = gsub(" ", "", response_final)) %>%
  with(table(water_direction_final, response_final))

# Run chi-square test
# Are there differences in cyanobacterial responses to increases vs. decreases in water level?
chisq_test(tab_cyano_all)
f_chisq_test(tab_cyano_all) #Bonferroni - adj. p

tab_cyano_down <- boxplotdata2 %>%
  mutate(
    water_direction_final = case_when(
      !is.na(cyano_response_wl_down) ~ "wl_down",
      is.na(cyano_response_wl_down) & !is.na(cyano_response_wl_up) ~ "wl_up",
      TRUE ~ NA_character_
    ),
    response_final = case_when(
      !is.na(cyano_response_wl_down) ~ cyano_response_wl_down,
      is.na(cyano_response_wl_down) & !is.na(cyano_response_wl_up) ~ cyano_response_wl_up,
      TRUE ~ NA_character_
    ),
    response_final = gsub(" ", "", response_final)
  ) %>%
  filter(!is.na(response_final)) %>%
  count(water_direction_final, response_final)

tab_cyano_down

tab_cyano_down <- xtabs(n ~ water_direction_final + response_final,
                    data = tab_cyano_down)

chisq_test(tab_cyano_down)
f_chisq_test(tab_cyano_down)
stats::fisher.test(tab_cyano_down)


tab_cyano_up <- boxplotdata2 %>%
  mutate(
    water_direction_final = case_when(
      !is.na(cyano_response_wl_up) ~ "wl_up",                     # priority
      is.na(cyano_response_wl_up) & !is.na(cyano_response_wl_down) ~ "wl_down",
      TRUE ~ NA_character_
    ),
    response_final = case_when(
      !is.na(cyano_response_wl_up) ~ cyano_response_wl_up,        # priority
      is.na(cyano_response_wl_up) & !is.na(cyano_response_wl_down) ~ cyano_response_wl_down,
      TRUE ~ NA_character_
    ),
    response_final = gsub(" ", "", response_final)
  ) %>%
  filter(!is.na(response_final)) %>%
  count(water_direction_final, response_final)


tab_cyano_up <- xtabs(n ~ water_direction_final + response_final,
                    data = tab_cyano_up)
chi_cyano_up <- stats::chisq.test(tab_cyano_up)
f_chisq_test(tab_cyano_up)


# Create contingency table
tab_phyto_all <- boxplotdata2 %>%
  pivot_longer(
    cols = c(phyto_response_wl_up, phyto_response_wl_down),
    names_to = "water_direction_final",
    values_to = "response_final"
  ) %>%
  mutate(response_final = gsub(" ", "", response_final)) %>%
  filter(!is.na(response_final)) %>%
  with(table(water_direction_final, response_final))

# Run chi-square test
# Are there differences in phytoplankton responses to increases vs. decreases in water level?
chisq_test(tab_phyto_all)
f_chisq_test(tab_phyto_all) #Bonferroni - adj. p

#wl decrease priority
tab_phyto <- boxplotdata2 %>%
  mutate(
    water_direction_final = case_when(
      !is.na(phyto_response_wl_down) ~ "wl_down",
      is.na(phyto_response_wl_down) & !is.na(phyto_response_wl_up) ~ "wl_up",
      TRUE ~ NA_character_
    ),
    response_final = case_when(
      !is.na(phyto_response_wl_down) ~ phyto_response_wl_down,
      is.na(phyto_response_wl_down) & !is.na(phyto_response_wl_up) ~ phyto_response_wl_up,
      TRUE ~ NA_character_
    ),
    response_final = gsub(" ", "", response_final)
  ) %>%
  filter(!is.na(response_final)) %>%
  count(water_direction_final, response_final)
tab_phyto
tab_phyto_decrease <- xtabs(n ~ water_direction_final + response_final,
                    data = tab_phyto)
tab_phyto_decrease

stats::chisq.test(tab_phyto_decrease)
f_chisq_test(tab_phyto_decrease)

tab_phyto_increase <- boxplotdata2 %>%
  mutate(
    water_direction_final = case_when(
      !is.na(phyto_response_wl_up) ~ "wl_up",                        # priority to wl_up
      is.na(phyto_response_wl_up) & !is.na(phyto_response_wl_down) ~ "wl_down",
      TRUE ~ NA_character_
    ),
    response_final = case_when(
      !is.na(phyto_response_wl_up) ~ phyto_response_wl_up,           # priority to wl_up
      is.na(phyto_response_wl_up) & !is.na(phyto_response_wl_down) ~ phyto_response_wl_down,
      TRUE ~ NA_character_
    ),
    response_final = gsub(" ", "", response_final)
  ) %>%
  filter(!is.na(response_final)) %>%
  count(water_direction_final, response_final)
tab_phyto_increase
phyto_increase_chi <- xtabs(n ~ water_direction_final + response_final,
                    data = tab_phyto_increase)
f_chisq_test(phyto_increase_chi)

stats::fisher.test(tab_cyano_down)
stats::fisher.test(tab_cyano_up)
stats::fisher.test(tab_phyto_decrease)
stats::fisher.test(phyto_increase_chi)
```

```{r}
tab <- tab_cyano_down
extract_chi_summary <- function(tab, model_name) {
  
  chi <- suppressWarnings(stats::chisq.test(tab))
  
  fisher_p <- tryCatch(
    stats::fisher.test(tab)$p.value,
    error = function(e) NA
  )

  
  tibble::tibble(
    model = model_name,
    n = sum(tab),
    chisq_stat = unname(chi$statistic),
    chisq_p = signif(chi$p.value, 3),
    fisher_p = signif(fisher_p, 3)
  )
}

summary_tbl_x2 <- dplyr::bind_rows(
  
  extract_chi_summary(tab_cyano_all, "cyano_all"),
  extract_chi_summary(tab_cyano_down, "cyano_down"),
  extract_chi_summary(tab_cyano_up, "cyano_up"),
  
  extract_chi_summary(tab_phyto_all, "phyto_all"),
  extract_chi_summary(tab_phyto_decrease, "phyto_decrease"),
  extract_chi_summary(phyto_increase_chi, "phyto_increase")
)

summary_tbl_x2


library(rstatix)
library(dplyr)
library(tidyr)
model_name = "cyano_all"
tab = tab_cyano_all
extract_cellwise_posthoc <- function(tab, model_name, p.adjust.method = "bonferroni") {


  # run global chi-square with descriptors
  chi_res <- tab %>% 
    chisq_test() %>%
    chisq_descriptives() %>%
    #std_residuals() %>%        # standardized residuals
    dplyr::select(water_direction_final, response_final, observed, expected, std.resid)

  # compute two‑sided p value from standardized residuals
  chi_res <- chi_res %>%
    mutate(p_raw = 2 * (1 - pnorm(abs(std.resid))))

  # multiple test correction
  chi_res <- chi_res %>%
    mutate(p_adj = p.adjust(p_raw, method = "bonferroni"),
           model = model_name)

  return(chi_res)
}

#create a summary table with all stats for Supplemental Table XX
summary_tbl <- dplyr::bind_rows(
  
   extract_cellwise_posthoc(tab_cyano_all, "cyano_all"),
    extract_cellwise_posthoc(tab_cyano_down, "cyano_down"),
  extract_cellwise_posthoc(tab_cyano_up, "cyano_up"),
  
  extract_cellwise_posthoc(tab_phyto_all, "phyto_all"),
  extract_cellwise_posthoc(tab_phyto_decrease, "phyto_decrease"),
  extract_cellwise_posthoc(phyto_increase_chi, "phyto_increase")
)   

summary_tbl_posthoc <- summary_tbl |> 
  mutate(significant = ifelse(p_adj < 0.05, "Significant", NA))

summary_tbl_posthoc
```

## Chi-squared test plots

```{r barplots for chisq}
library(dplyr)
cyano_chi <- ggplot(as.data.frame(tab_cyano_all), aes(x = water_direction_final, y = Freq, fill = response_final)) +
  geom_bar(stat = "identity", position = "fill") +
     geom_text(aes(label = Freq),
            position = position_fill(vjust = 0.5), 
            color = "black", 
            size = 6) +
  labs(x = "",
       y = "Proportion",
       fill = "Response") + #,
      # title = "Cyanobacteria Response by Water Level Direction") +
  scale_x_discrete(labels = c(
  "cyano_response_wl_up" = "Water level increase",
  "cyano_response_wl_down" = "Water level decrease"))+
 scale_fill_manual(
  values = c("#b2ebf2", "#4dd0e1", "#e0e0e0"),   # 3 colors
  name = "Did \ncyanobacteria\nincrease or \ndecrease?",
  labels = c("Decrease", "Increase", "No Change")) +
  theme_bw() + 
    theme(axis.text.x = element_text(hjust = 0.5, size = 12),
        #axis.text.x = element_blank(),
        axis.text.y = element_text(hjust = 1, size = 12),
        axis.title.y=element_text(angle=90, vjust = 0.5, size = 12),
        axis.title.x=element_text(vjust = 1, size = 12),
        #legend.text  = element_blank(),
        legend.title=element_text(size = 12), 
        legend.key.height = unit(1, 'cm'),
        legend.key.width = unit(1, 'cm')) 

df_plot_phyto <- as.data.frame(tab_phyto_all)

phyto_chi <- ggplot(df_plot_phyto, aes(x = water_direction_final, y = Freq, fill = response_final)) +
  geom_bar(stat = "identity", position = "fill") +
     geom_text(aes(label = Freq),
            position = position_fill(vjust = 0.5), 
            color = "black", 
            size = 6) +
  labs(x = "",
       y = "Proportion",
       fill = "Response") +#,
      # title = "Phytoplankton Response by Water Level Direction") +
    scale_x_discrete(labels = c(
  "phyto_response_wl_up" = "Water level increase",
  "phyto_response_wl_down" = "Water level decrease"))+
  scale_fill_manual(values = c("#ddefcd", "#99bbab", "#e0e0e0"), 
                    name = "Did \nphytoplankton\nincrease or \ndecrease?",
  labels = c("Decrease", "Increase", "No Change")) +
  theme_bw() + 
    theme(axis.text.x = element_text(hjust = 0.5, size = 12),
        #axis.text.x = element_blank(),
        axis.text.y = element_text(hjust = 1, size = 12),
        axis.title.y=element_text(angle=90, vjust = 0.5, size = 12),
        axis.title.x=element_text(vjust = 1, size = 12),
        #legend.text  = element_blank(),
        legend.title=element_text(size = 12), 
        legend.key.height = unit(1, 'cm'),
        legend.key.width = unit(1, 'cm')) 

chi <- phyto_chi / cyano_chi + #+ plot_annotation(title = "Chi-Square") +
  plot_layout(axes = "collect")


chi

ggsave(paste0(here(), "/Manuscript_figures/chi.jpg"), plot = chi, height = 5, width = 6.5, units = "in")
```

# Supplemental figures

```{r create a folder for supplemental figures}
supplement_figures <- here("Supplement_figures/")
if(!file.exists(supplement_figures)){
    #make the folder
    dir.create(file.path(supplement_figures))
  }
```

## Trophic Levels

```{r trophic levels}
library(dplyr)
#install.packages("conflicted")
library(conflicted)
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")
#consider subsetting these to phyto and cyano specific studies?
non_dups <- data |> select(res_name, trophic_status_reduced) |> 
  unique()

#plot only reduced trophic state levels and only reservoirs that are not duplicated
#from 81 -> 72
trophic <- non_dups |> 
  filter(trophic_status_reduced != "not reported") |> 
  ggplot(aes(x= factor(trophic_status_reduced, c("oligotrophic", "mesotrophic", "eutrophic", "hypereutrophic")))) + 
  geom_bar(fill = "#608687") +
  xlab("Trophic Status") +
  ylab("Number of reservoirs") + 
  theme_bw() +
  #theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))+
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, size = 5, colour = "black")

trophic

ggsave(paste0(here(),"/Supplement_figures/trophic.jpg"), plot = trophic, height = 3, width = 6.5, units = "in")

```

## Logistic regression prioritizing increases in water level

Logistic regression figures that prioritize either increases or decreases in water level, rather than both as plotted above.

```{r logreg increase priority}
cyanodecreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_priorityincrease, combined_cyano_response_increasepriority, combined_cyano_response_binary_increasepriority, name_yr) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>%
  filter(!is.na(combined_cyano_response_binary_increasepriority)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_binaryincrease = combined_cyano_response_binary_increasepriority, 
         response_increase = combined_cyano_response_increasepriority) %>% 
  mutate(type = "cyano")

cyanoincreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_priorityincrease, combined_cyano_response_increasepriority, combined_cyano_response_binary_increasepriority, name_yr) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>%
  filter(!is.na(combined_cyano_response_binary_increasepriority)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_binaryincrease = combined_cyano_response_binary_increasepriority, 
         response_increase = combined_cyano_response_increasepriority) %>% 
  mutate(type = "cyano")

phytoincreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_priorityincrease, combined_phyto_response_increasepriority, combined_phyto_response_binary_increasepriority, name_yr) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>% 
  filter(!is.na(combined_phyto_response_binary_increasepriority)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_binaryincrease = combined_phyto_response_binary_increasepriority, 
         response_increase = combined_phyto_response_increasepriority) %>% 
  mutate(type = "phyto")

phytodecreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_priorityincrease, combined_phyto_response_increasepriority, combined_phyto_response_binary_increasepriority, name_yr) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>% 
  filter(!is.na(combined_phyto_response_binary_increasepriority)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_binaryincrease = combined_phyto_response_binary_increasepriority, 
         response_increase = combined_phyto_response_increasepriority) %>% 
  mutate(type = "phyto")
  
cyanodata <- boxplotdata2 %>% 
  select(prop_delta_wl_priorityincrease, combined_cyano_response_increasepriority, combined_cyano_response_binary_increasepriority, name_yr, trophic_status_reduced) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>%
  filter(!is.na(combined_cyano_response_binary_increasepriority)) %>% 
  # filter(prop_delta_wl<0) %>% 
  rename(response_binary = combined_cyano_response_binary_increasepriority, 
         response = combined_cyano_response_increasepriority) %>% 
  mutate(type = "cyano") |> 
  mutate(wl_priority = "increase")

phytodata <- boxplotdata2 %>% 
  select(prop_delta_wl_priorityincrease, combined_phyto_response_increasepriority, combined_phyto_response_binary_increasepriority, name_yr, trophic_status_reduced) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>% 
  filter(!is.na(combined_phyto_response_binary_increasepriority)) %>% 
  # filter(prop_delta_wl<0) %>% 
  rename(response_binary = combined_phyto_response_binary_increasepriority, 
         response = combined_phyto_response_increasepriority) %>% 
  mutate(type = "phyto")|> 
  mutate(wl_priority = "increase")
# logregdata <- rbind(cyanoincreasedata, cyanodecreasedata, phytoincreasedata, phytodecreasedata)
logregdata_increase <- rbind(cyanodata, phytodata)
#"#7A524B","#3C7F29","#8DBAE9"

logreg_wlincrease <- logregdata_increase %>% ggplot(aes(x = prop_delta_wl_priorityincrease, y = response_binary, color = as.factor(type), shape = as.factor(type))) +
  scale_color_manual(values = c("#88CCEE", "#117733"), name = "")+
  scale_shape_manual(values = c(8, 20)) +
  geom_point(size = 3) +
  geom_smooth(method = "glm", 
              method.args = list(family = "binomial"),
              se = FALSE, linewidth = 3) + 
  #geom_jitter()+
 ylab("Response to changes \nin water level") + 
  xlab("Proportion water level change") + 
  ggtitle("Priority WL increases") + 
  theme_bw() +
  theme(axis.text.x = element_text(hjust = 1, size = 10), 
        axis.title.y=element_text(angle=90, vjust = .5, size = 12), 
        legend.title=element_blank())
logreg_wlincrease

logregdata_phyto <- logregdata_increase %>% subset(type == "phyto")
logregdata_cyano <- logregdata_increase %>% subset(type == "cyano")

logregform_phyto <-  glm(formula=response_binary ~ prop_delta_wl_priorityincrease, data=logregdata_phyto, family=binomial)
summary(logregform_phyto)

logregform_cyano <-  glm(formula=response_binary ~ prop_delta_wl_priorityincrease, data=logregdata_cyano, family=binomial)
summary(logregform_cyano)
ggsave(paste0(here(),"/Supplement_figures/logreg_wlincrease.jpg"), plot = logreg_wlincrease, height = 3, width = 6.5, units = "in")
```

## Logistic regression prioritizing decreases in water level

```{r logreg wl decrease priority}
cyanodecreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_prioritydecrease, combined_cyano_response, combined_cyano_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl_prioritydecrease)) %>%
  filter(!is.na(combined_cyano_response_binary)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_binary = combined_cyano_response_binary, 
         response = combined_cyano_response) %>% 
  mutate(type = "cyano")

cyanoincreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_prioritydecrease, combined_cyano_response, combined_cyano_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl_prioritydecrease)) %>%
  filter(!is.na(combined_cyano_response_binary)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_binary = combined_cyano_response_binary, 
         response = combined_cyano_response) %>% 
  mutate(type = "cyano")

phytoincreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_prioritydecrease, combined_phyto_response, combined_phyto_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl_prioritydecrease)) %>% 
  filter(!is.na(combined_phyto_response_binary)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_binary = combined_phyto_response_binary, 
         response = combined_phyto_response) %>% 
  mutate(type = "phyto")

phytodecreasedata <- boxplotdata2 %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_prioritydecrease, combined_phyto_response, combined_phyto_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl_prioritydecrease)) %>% 
  filter(!is.na(combined_phyto_response_binary)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_binary = combined_phyto_response_binary, 
         response = combined_phyto_response) %>% 
  mutate(type = "phyto")
  
cyanodata <- boxplotdata2 %>% 
  select(prop_delta_wl_prioritydecrease, combined_cyano_response, combined_cyano_response_binary, name_yr, trophic_status_reduced) %>% 
  filter(!is.na(prop_delta_wl_prioritydecrease)) %>%
  filter(!is.na(combined_cyano_response_binary)) %>% 
  # filter(prop_delta_wl_prioritydecrease,<0) %>% 
  rename(response_binary = combined_cyano_response_binary, 
         response = combined_cyano_response) %>% 
  mutate(type = "cyano") |> 
  mutate(wl_priority = "decrease")

phytodata <- boxplotdata2 %>% 
  select(prop_delta_wl_prioritydecrease, combined_phyto_response, combined_phyto_response_binary, name_yr, trophic_status_reduced) %>% 
  filter(!is.na(prop_delta_wl_prioritydecrease)) %>% 
  filter(!is.na(combined_phyto_response_binary)) %>% 
  # filter(prop_delta_wl_prioritydecrease,<0) %>% 
  rename(response_binary = combined_phyto_response_binary, 
         response = combined_phyto_response) %>% 
  mutate(type = "phyto") |> 
  mutate(wl_priority = "decrease")

# logregdata <- rbind(cyanoincreasedata, cyanodecreasedata, phytoincreasedata, phytodecreasedata)
logregdata_decrease <- rbind(cyanodata, phytodata)
logregdata <- logregdata_decrease
#"#7A524B","#3C7F29","#8DBAE9"

logreg_wldecrease <- logregdata %>% ggplot(aes(x = prop_delta_wl_prioritydecrease, y = response_binary, color = as.factor(type), shape = as.factor(type))) +
  scale_color_manual(values = c("#88CCEE", "#117733"), name = "")+
  scale_shape_manual(values = c(8, 20)) +
  geom_point(size = 3) +
  geom_smooth(method = "glm", 
              method.args = list(family = "binomial"),
              se = FALSE, linewidth = 3) + 
  #geom_jitter()+
 ylab("Response to changes \nin water level") + 
  xlab("Proportion water level change") + 
  ggtitle("Priority WL decrease")+
  theme_bw() +
  theme(axis.text.x = element_text(hjust = 1, size = 10), 
        axis.title.y=element_text(angle=90, vjust = .5, size = 12), 
        legend.title=element_blank())
logreg_wldecrease

logregdata_phyto <- logregdata %>% subset(type == "phyto")
logregdata_cyano <- logregdata %>% subset(type == "cyano")

logregform_phyto <-  glm(formula=response_binary ~ prop_delta_wl_prioritydecrease, data=logregdata_phyto, family=binomial)
summary(logregform_phyto)

logregform_cyano <-  glm(formula=response_binary ~ prop_delta_wl_prioritydecrease, data=logregdata_cyano, family=binomial)
summary(logregform_cyano)
both_logreg <- logreg_wldecrease/logreg_wlincrease
ggsave(paste0(here(),"/Supplement_figures/logreg_wldecrease.jpg"), plot = logreg_wlincrease, height = 3, width = 6.5, units = "in")
ggsave(paste0(here(),"/Supplement_figures/bothlogreg.jpg"), plot = both_logreg, height = 6, width = 6.5, units = "in")
```

## Barplot: number of reservoir by continent

```{r reservoir by continent}

#countries by continent
countries_sf <- data |> 
  select(lat_dd, long_dd, res_name, cov_num, name_yr) |> 
  filter(!is.na(lat_dd) & !is.na(long_dd)) |> 
  st_as_sf(coords = c("long_dd", "lat_dd"), crs = 4326) 
world <- ne_countries(scale = "medium", returnclass = "sf")
df_with_country <- st_join(countries_sf, world["name_long"])
df_final <- df_with_country %>% 
  st_drop_geometry() %>%
  rename(country = name_long)
df_final <- df_final %>% 
  mutate(continent = world$continent[match(country, world$name_long)])

countries <- left_join(data, df_final, by = c("res_name", "cov_num", "name_yr")) |> select(c("res_name", "name_yr", "country.x", "country.y", "continent")) |> 
  mutate(country.x = ifelse(country.x == "", NA, country.x)) |> 
  mutate(country = ifelse(!is.na(country.x), country.x, country.y)) |> 
  mutate(continent.a = case_when(
    country %in% c("Brazil", "Argentina", "Uruguay", "Argentina and Uruguay") ~ "South America",
      country %in% c("United States", "USA/Canada", "Canada", "Mexico") ~ "North America",
      country %in% c("Portugal", "England", "Italy", "France", "Serbia", 
                     "Estonia", "Estonia/Russia") ~ "Europe",
      country %in% c("Russia") ~ "Europe/Asia",
      country %in% c("China", "Taiwan", "Malaysia", "Mongolia", "Lebanon",
                     "Cambodia", "Vietnam") ~ "Asia",
       country %in% c("Australia") ~ "Oceania",
    country %in% c("South Africa") ~ "Africa",
      TRUE ~ NA_character_)) |> 
  mutate(continent = ifelse(!is.na(continent.a), continent.a, continent)) |> 
  select(res_name, country, continent) |> 
  unique()

countries |> 
  filter(!is.na(continent)) |> 
  ggplot(aes(x=continent))+
  geom_bar() + 
  xlab("Continent") +
  ylab("Number of reservoirs") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))+
  geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, size = 5, colour = "black")
```

## Histogram of maximum reservoir depths and surface areas

```{r maximum depths}
#histogram of maxdepths

maxdepth <- data |> 
  subset(!is.na(fullpond_maxdepth)) %>% 
  ggplot(aes(x=fullpond_maxdepth)) + 
  geom_histogram() +
  xlab("Max depth at full pond (m)") +
  ylab("Number of lakes") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggsave(paste0(here(),"/Supplement_figures/maxdepth.jpg"), plot = maxdepth, height = 3, width = 6.5, units = "in")

maxsa <- data |> 
  subset(!is.na(fullpond_sa)) %>% 
  ggplot(aes(x=fullpond_sa)) + 
  geom_histogram() +
  xlab("Max surface area at full pond (km2)") +
  ylab("Number of lakes") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

maxsadepth <- maxdepth/maxsa + plot_layout(axes = "collect")

ggsave(paste0(here(),"/Supplement_figures/maxsa.jpg"), plot = maxsa, height = 3, width = 6.5, units = "in")

ggsave(paste0(here(),"/Supplement_figures/maxcombined.jpg"), plot = maxsadepth, height = 3, width = 4, units = "in")
```

## Drivers for phyto + cyano combined

drivers and look at variables that changed with WL change

```{r combined drivers}
drivers <- data |> 
  mutate(tp = ifelse(grepl("TP", chem_drive), 1, 0),                            #chemical variables
         tn = ifelse(grepl("TN", chem_drive), 1, 0), 
         nh4 = ifelse(grepl("NH4", chem_drive), 1, 0), 
         ph = ifelse(grepl("pH", chem_drive), 1, 0),
         do = ifelse(grepl("DO", chem_drive), 1, 0),
         srp = ifelse(grepl("SRP", chem_drive), 1, 0),
         no3 = ifelse(grepl("NO3", chem_drive), 1, 0),
         cond = ifelse(grepl("conductivity", chem_drive), 1, 0),
         metals = ifelse(grepl("metals", chem_drive), 1, 0),
         orp = ifelse(grepl("ORP", chem_drive), 1, 0),
         no2 = ifelse(grepl("NO2", chem_drive), 1, 0),
         dic = ifelse(grepl("DIC", chem_drive), 1, 0),
       dom = ifelse(grepl("DOM", chem_drive), 1, 0),
       pom = ifelse(grepl("POM", chem_drive), 1, 0),
       tkn = ifelse(grepl("TKN", chem_drive), 1, 0),
         doc = ifelse(grepl("DOC", chem_drive), 1, 0),
         din = ifelse(grepl("DIN", chem_drive), 1, 0),
         ph = ifelse(grepl("pH", chem_drive), 1, 0),
         salinity = ifelse(grepl("salinity", chem_drive), 1, 0), 
         other_chem = ifelse(grepl("Other", chem_drive), 1, 0)) |> 
  mutate(water_level = ifelse(grepl("water level", phys_drive), 1, 0), 
         water_temp = ifelse(grepl("water temperature", phys_drive), 1, 0),
         transparency = ifelse(grepl("transparency", phys_drive), 1, 0),
         turb = ifelse(grepl("turbidity", phys_drive), 1, 0),
         light = ifelse(grepl("PAR", phys_drive), 1, 0),    #need to pull out rainfall & precipitation, stability/, water residence time, flushing
         tss_tds = ifelse(grepl("total suspended solids|tss|tds|total dissolved solid", phys_drive), 1, 0), 
         strat = ifelse(grepl("mix|stab|strat", phys_drive), 1, 0),
         restime = ifelse(grepl("residence", phys_drive), 1, 0), 
         flush_discharge = ifelse(grepl("flush|discharge", phys_drive), 1, 0), 
         inflow = ifelse(grepl("inflow", phys_drive), 1, 0),
         precip = ifelse(grepl("precip|rain", phys_drive), 1, 0),
         air_temp = ifelse(grepl("air", phys_drive), 1, 0),
         other_phys = ifelse(grepl("Other", phys_drive), 1, 0))

other_phys <- drivers |> 
  filter(other_phys == 1) |> 
  select(name_yr, title, num_water, phys_drive)

  unique(data$phys_response)
  
  #WLchangevars is whether or not vars changed under WL change
WLchangevars <- data |> 
mutate(not_assessed = ifelse(grepl("not assessed", phys_response), 1, 0), 
         water_temp = ifelse(grepl("water temperature", phys_response), 1, 0),
         transparency = ifelse(grepl("transparency", phys_response), 1, 0),
         turb = ifelse(grepl("turbidity", phys_response), 1, 0),
         light = ifelse(grepl("PAR", phys_response), 1, 0),    #need to pull out rainfall & precipitation, stability/, water residence time, flushing
         tss_tds = ifelse(grepl("total suspended solids|tss|tds|total dissolved solid|sediment resuspension", phys_response), 1, 0), 
         #strat = ifelse(grepl("mix|stab|strat", phys_response), 1, 0),
         #restime = ifelse(grepl("residence", phys_response), 1, 0), 
         #flush_discharge = ifelse(grepl("flush|discharge", phys_response), 1, 0), 
         #inflow = ifelse(grepl("inflow", phys_response), 1, 0),
         #precip = ifelse(grepl("precip|rain", phys_response), 1, 0),
         #air_temp = ifelse(grepl("air", phys_response), 1, 0),
         other_phys = ifelse(grepl("Other", phys_response), 1, 0), 
       watercolor = ifelse(grepl("water color", phys_response), 1, 0)) |> 
mutate(tp = ifelse(grepl("TP", chem_response), 1, 0),                            #chemical variables
         tn = ifelse(grepl("TN", chem_response), 1, 0), 
         nh4 = ifelse(grepl("NH4", chem_response), 1, 0), 
         ph = ifelse(grepl("pH", chem_response), 1, 0),
         do = ifelse(grepl("DO", chem_response), 1, 0),
         srp = ifelse(grepl("SRP", chem_response), 1, 0),
         no3 = ifelse(grepl("NO3", chem_response), 1, 0),
         cond = ifelse(grepl("conductivity", chem_response), 1, 0),
         metals = ifelse(grepl("metals", chem_response), 1, 0),
         orp = ifelse(grepl("ORP", chem_response), 1, 0),
         no2 = ifelse(grepl("NO2", chem_response), 1, 0),
         doc = ifelse(grepl("DOC", chem_response), 1, 0),
       dic = ifelse(grepl("DIC", chem_response), 1, 0),
       dom = ifelse(grepl("DOM", chem_response), 1, 0),
       pom = ifelse(grepl("POM", chem_response), 1, 0),
       tkn = ifelse(grepl("TKN", chem_response), 1, 0),
         din = ifelse(grepl("DIN", chem_response), 1, 0),
         ph = ifelse(grepl("pH", chem_response), 1, 0),
         salinity = ifelse(grepl("salinity", chem_response), 1, 0), 
         other_chem = ifelse(grepl("Other", chem_response), 1, 0)) 

#spreadvars is whether or not the var was measured
spreadvars <- data |> 
  mutate(filt_chla = ifelse(grepl("filtered chlorophyll", bio_vars), 1, 0),                       #biotic variables
         phyto_micros = ifelse(grepl("phytoplankton microscopy", bio_vars), 1, 0), 
         single_cyano_taxon = ifelse(grepl("single cyanobacterial taxon", bio_vars), 1, 0), 
         sensor_chla = ifelse(grepl("sensor", bio_vars), 1, 0),
         other_bio = ifelse(grepl("Other", bio_vars), 1, 0)) |> 
  mutate(tp = ifelse(grepl("TP", chem_vars), 1, 0),                            #chemical variables
         tn = ifelse(grepl("TN", chem_vars), 1, 0), 
         nh4 = ifelse(grepl("NH4", chem_vars), 1, 0), 
         ph = ifelse(grepl("pH", chem_vars), 1, 0),
         do = ifelse(grepl("DO", chem_vars), 1, 0),
         srp = ifelse(grepl("SRP", chem_vars), 1, 0),
         no3 = ifelse(grepl("NO3", chem_vars), 1, 0),
         cond = ifelse(grepl("conductivity", chem_vars), 1, 0),
         metals = ifelse(grepl("metals", chem_vars), 1, 0),
         orp = ifelse(grepl("ORP", chem_vars), 1, 0),
         no2 = ifelse(grepl("NO2", chem_vars), 1, 0),
         doc = ifelse(grepl("DOC", chem_vars), 1, 0),
          dic = ifelse(grepl("DIC", chem_vars), 1, 0),
       dom = ifelse(grepl("DOM", chem_vars), 1, 0),
       pom = ifelse(grepl("POM", chem_vars), 1, 0),
       tkn = ifelse(grepl("TKN", chem_vars), 1, 0),
         din = ifelse(grepl("DIN", chem_vars), 1, 0),
         ph = ifelse(grepl("pH", chem_vars), 1, 0),
         salinity = ifelse(grepl("salinity", chem_vars), 1, 0), 
         other_chem = ifelse(grepl("Other", chem_vars), 1, 0)) |> 
  mutate(water_level = ifelse(grepl("water level", phys_vars), 1, 0), 
         water_temp = ifelse(grepl("water temperature", phys_vars), 1, 0),
         transparency = ifelse(grepl("transparency", phys_vars), 1, 0),
         turb = ifelse(grepl("turbidity", phys_vars), 1, 0),
         light = ifelse(grepl("PAR", phys_vars), 1, 0),    #need to pull out rainfall & precipitation, stability/, water residence time, flushing #these are commented out below
        tss_tds = ifelse(grepl("total suspended solids|tss|tds|total dissolved solid|sediment resuspension", phys_vars), 1, 0), 
         #strat = ifelse(grepl("mix|stab|strat", phys_vars), 1, 0),
        # restime = ifelse(grepl("residence", phys_vars), 1, 0), 
        # flush_discharge = ifelse(grepl("flush|discharge", phys_vars), 1, 0), 
        # inflow = ifelse(grepl("inflow", phys_vars), 1, 0),
        # precip = ifelse(grepl("precip|rain", phys_vars), 1, 0),
        # air_temp = ifelse(grepl("air", phys_vars), 1, 0),
         other_phys = ifelse(grepl("Other", phys_vars), 1, 0)) 
        # watercolor = ifelse(grepl("water color", phys_vars), 1, 0))

# unique(data$chem_drive)
# unique(data$phys_drive)

#drivers
chemvars_drive <- drivers |> select(c(cond, do, nh4, no2, no3, orp, ph, srp, tn, tp))
freqchem_drive <- as.data.frame(colSums(chemvars_drive))
chemvars_drive <- setNames(cbind(rownames(freqchem_drive), freqchem_drive, row.names = NULL), 
         c("names", "sum"))

physvars_drive <- drivers |> select(c(water_level, water_temp, transparency, turb, light, tss_tds, other_phys))
freqphys_drive <- as.data.frame(colSums(physvars_drive))
physvars_drive <- setNames(cbind(rownames(freqphys_drive), freqphys_drive, row.names = NULL), 
         c("names", "sum"))

#all measured
chemvars <- spreadvars |> select(c(cond, do, nh4, no2, no3, orp, ph, srp, tn, tp))
freqchem <- as.data.frame(colSums(chemvars))
chemvars <- setNames(cbind(rownames(freqchem), freqchem, row.names = NULL), 
         c("names", "measured"))

physvars <- spreadvars |> select(c(water_level, water_temp, transparency, turb, light, tss_tds, other_phys))
freqphys <- as.data.frame(colSums(physvars))
physvars <- setNames(cbind(rownames(freqphys), freqphys, row.names = NULL), 
         c("names", "measured"))

physvars_noWL <- spreadvars |> select(c(water_temp, transparency, turb, light, tss_tds, other_phys))
freqphys_noWL<- as.data.frame(colSums(physvars_noWL))
physvars_noWL <- setNames(cbind(rownames(freqphys_noWL), freqphys_noWL, row.names = NULL), 
         c("names", "measured"))

#changes with WL change
chemvarsWL <- WLchangevars |> select(c(cond, do, nh4, no2, no3, orp, ph, srp, tn, tp))
freqchemWL <- as.data.frame(colSums(chemvarsWL))
chemvarsWL <- setNames(cbind(rownames(freqchemWL), freqchemWL, row.names = NULL), 
         c("names", "sum"))

physvarsWL <- WLchangevars |> select(c(water_temp, transparency, turb, light, tss_tds, other_phys))
freqphysWL <- as.data.frame(colSums(physvarsWL))
physvarsWL <- setNames(cbind(rownames(freqphysWL), freqphysWL, row.names = NULL), 
         c("names", "sum"))

#physvarsWL_all <- WLchangevars |> select(c(water_temp, transparency, turb, light, tss_tds, other_phys, strat, restime, flush_discharge, inflow, precip, air_temp)) # this is all possible variables, our analysis was limited to only the ones we specifically extracted

chemvars_drive_comb <-  right_join(chemvars, chemvars_drive)
physvars_drive_comb <- right_join(physvars, physvars_drive)

chemvarsWL_comb <- right_join(chemvarsWL, chemvars, by = "names")
physvarsWL_comb <- right_join(physvars_noWL, physvarsWL, by = "names")

#Make supplemental Figs
#water level change
chemWL <- chemvarsWL_comb |> 
  filter(!names == "other_chem") |> 
  select(c(names, measured, sum)) |> 
  rename(id = sum) |> 
  mutate(notid = measured - id,
         prop_id = id / measured) |>            # calculate proportion
  select(c(names, notid, id, prop_id)) |> 
  arrange(desc(prop_id)) |>                     # order by highest proportion

  mutate(names = factor(names, levels = names)) |> 
  select(-prop_id) |>                           # remove helper column
  pivot_longer(!names, names_to = "driver", values_to = "count") |> 
  ggplot(aes(fill = driver, y = count, x = names)) +
  geom_bar(position = position_fill(reverse = TRUE),  stat = "identity") + 
  #ggtitle("Chemical Variables") + 
  geom_text(aes(label = count),
            position = position_fill(vjust = 0.5, reverse = TRUE), 
            color = "black", 
            size = 3) + 
  scale_fill_manual(values = c("#8680cc", "#cfcdef"),
  labels = c("Yes", "No"), 
  name = "Did authors \nidentify this \nvariable as changing \nwith water level?") +
  theme_bw() + 
  labs(y = "") +
  theme(axis.text.x  = element_text(hjust = .5, size = 6),
        axis.title.x = element_blank(),
        axis.title.y = element_text(angle = 90, vjust = .5, size = 12))

physWL <- physvarsWL_comb |> 
  filter(!names == "other_phys") |> 
  select(c(names, measured, sum)) |> 
  rename(id = sum) |> 
  mutate(notid = measured - id,
         prop_id = id / measured) |>            # calculate proportion
  select(c(names, notid, id, prop_id)) |> 
  arrange(desc(prop_id)) |>                     # order by highest proportion

  mutate(names = factor(names, levels = names)) |> 
  select(-prop_id) |>                           # remove helper column
  pivot_longer(!names, names_to = "driver", values_to = "count") |> 
  ggplot(aes(fill = driver, y = count, x = names)) +
  geom_bar(position = position_fill(reverse = TRUE),  stat = "identity") + 
  ggtitle("") + 
  geom_text(aes(label = count),
            position = position_fill(vjust = 0.5, reverse = TRUE), 
            color = "black", 
            size = 3) + 
  scale_fill_manual(values = c("#8680cc", "#cfcdef"),
  labels = c("Yes", "No"), 
  name = "Did authors \nidentify this \nvariable as changing \nwith water level?") +
  theme_bw() + 
  labs(y = "") +
  theme(axis.text.x  = element_text(hjust = .5, size = 6),
        axis.title.x = element_blank(),
        axis.title.y = element_text(angle = 90, vjust = .5, size = 12))


chemWL <- chemvarsWL_comb |> 
  filter(!names == "other_chem") |> 
  select(c(names, measured, sum)) |> 
  rename(id = sum) |> 
  mutate(notid = measured - id,
         prop_id = id / measured) |>            # calculate proportion
  select(c(names, notid, id, prop_id)) |> 
  arrange(desc(prop_id)) |>                     # order by highest proportion

  mutate(names = factor(names, levels = names)) |> 
  select(-prop_id) |>                           # remove helper column
  pivot_longer(!names, names_to = "driver", values_to = "count") |> 
  ggplot(aes(fill = driver, y = count, x = names)) +
  geom_bar(position = position_fill(reverse = TRUE),  stat = "identity") + 
  ggtitle("") + 
  geom_text(aes(label = count),
            position = position_fill(vjust = 0.5, reverse = TRUE), 
            color = "black", 
            size = 3) + 
  scale_fill_manual(values = c("#8680cc", "#cfcdef"),
  labels = c("Yes", "No"), 
  name = "Did authors \nidentify this \nvariable as changing \nwith water level?") +
  theme_bw() + 
  labs(y = "") +
  theme(axis.text.x  = element_text(hjust = .5, size = 6),
        axis.title.x = element_blank(),
        axis.title.y = element_text(angle = 90, vjust = .5, size = 12))

#overall response
physdrive <- physvars_drive_comb |> 
 filter(!names == "other_phys") |> 
  select(c(names, measured, sum)) |> 
  rename(id = sum) |> 
  mutate(measured = if_else(measured == 42, 61, measured),
         notid = measured - id,
         prop_id = id / measured) |>            # calculate proportion
  select(c(names, notid, id, prop_id)) |> 
  arrange(desc(prop_id)) |>                     # order by highest proportion

  mutate(names = factor(names, levels = names)) |> 
  select(-prop_id) |>                           # remove helper column
  pivot_longer(!names, names_to = "driver", values_to = "count") |> 
  ggplot(aes(fill = driver, y = count, x = names)) +
  geom_bar(position = position_fill(reverse = TRUE),  stat = "identity") + 
  ggtitle("Physical Variables") + 
  geom_text(aes(label = count),
            position = position_fill(vjust = 0.5, reverse = TRUE), 
            color = "black", 
            size = 3) + 
  scale_fill_manual(values = c("#CC6060", "#F2DADA"), 
  labels = c("Yes", "No"), 
  name = "Did authors \nidentify this \nvariable as driving \nresponse?") +
  theme_bw() + 
  labs(y = "") +
    theme(#legend.position = "none",
  axis.text.x  = element_text(hjust = .5, size = 6),
  axis.title.x = element_blank(),
  axis.title.y = element_text(angle = 90, vjust = .5, size = 12))

chemdrive <- chemvars_drive_comb |> 
  filter(!names == "other_chem") |> 
  select(c(names, measured, sum)) |> 
  rename(id = sum) |> 
  mutate(notid = measured - id,
         prop_id = id / measured) |>            # calculate proportion
  select(c(names, notid, id, prop_id)) |> 
  arrange(desc(prop_id)) |>                     # order by highest proportion

  mutate(names = factor(names, levels = names)) |> 
  select(-prop_id) |>                           # remove helper column
  pivot_longer(!names, names_to = "driver", values_to = "count") |> 
  ggplot(aes(fill = driver, y = count, x = names)) +
  geom_bar(position = position_fill(reverse = TRUE),  stat = "identity") + 
  ggtitle("Chemical Variables") + 
  geom_text(aes(label = count),
            position = position_fill(vjust = 0.5, reverse = TRUE), 
            color = "black", 
            size = 3) + 
  scale_fill_manual(values = c("#CC6060", "#F2DADA"),
  labels = c("Yes", "No"), 
  name = "Did authors \nidentify this \nvariable as driving \nresponse?") +
  theme_bw() + 
  labs(y = "") +
  theme(#legend.position = "none",
  axis.text.x  = element_text(hjust = .5, size = 6),
  axis.title.x = element_blank(),
  axis.title.y = element_text(angle = 90, vjust = .5, size = 12))

combined <- ((chemdrive/physdrive) | (chemWL / physWL)) + plot_layout(guides = "collect") &
  theme(
    legend.position = "right",
    legend.box = "vertical"
  )
combined

ggsave(paste0(here(), "/Supplement_figures/drivers_all_WL.jpg"), plot = combined, height = 4, width = 8, units = "in")
```

## Reverse Logistic Regression

```{r reverse logreg}

boxplotdata2_opposite <- boxplotdata2 %>% 
  #mutate(prop_delta_wl = delta_wl_direction/depth_max_m) %>% 
    mutate(phyto_binary_increase = ifelse(increase_phyto_response == "No significant response", 1, 
                                ifelse(increase_phyto_response == "Decrease in phytoplankton", 0, 
                                ifelse(increase_phyto_response == "Increase in phytoplankton", 1, NA)))) %>% 
  mutate(phyto_binary_decrease = ifelse(decrease_phyto_response == "No significant response", 1, 
                                ifelse(decrease_phyto_response == "Decrease in phytoplankton", 0, 
                                ifelse(decrease_phyto_response == "Increase in phytoplankton", 1, NA)))) %>%
  # mutate(combined_phyto_response_text = decrease_phyto_response %>% 
  #          is.na() %>% 
  #          ifelse(increase_phyto_response, decrease_phyto_response)) %>% 
    mutate(combined_phyto_response_binary = case_when(
  is.na(phyto_binary_decrease) & prop_delta_wl_prioritydecrease < 0 ~ NA_real_,
  is.na(phyto_binary_decrease) ~ phyto_binary_increase,
  TRUE ~ phyto_binary_decrease
)) |> 
    mutate(combined_phyto_response_binary_increasepriority = case_when(
 prop_delta_wl_priorityincrease < 0 ~ phyto_binary_decrease,
  is.na(phyto_binary_increase) & prop_delta_wl_priorityincrease > 0 ~ phyto_binary_increase,
  TRUE ~ phyto_binary_increase
)) |> 
    #mutate(combined_phyto_response_binary = ifelse(is.na(phyto_binary_decrease), phyto_binary_increase, phyto_binary_decrease)) %>% 
    mutate(cyano_binary_increase = ifelse(increase_cyano_response == "No significant response", 0, 
                                ifelse(increase_cyano_response == "Decrease in cyanobacteria", 0, 
                                ifelse(increase_cyano_response == "Increase in cyanobacteria", 1, NA)))) %>% 
  mutate(cyano_binary_decrease = ifelse(decrease_cyano_response == "No significant response", 0, 
                                ifelse(decrease_cyano_response == "Decrease in cyanobacteria", 0, 
                                ifelse(decrease_cyano_response == "Increase in cyanobacteria", 1, NA)))) %>%
  # mutate(combined_phyto_response_text = decrease_phyto_response %>% 
  #          is.na() %>% 
  #          ifelse(increase_phyto_response, decrease_phyto_response)) %>% 
    #mutate(combined_cyano_response_binary = ifelse(is.na(cyano_binary_decrease), cyano_binary_increase, cyano_binary_decrease)) %>% 
  mutate(cyano_response_wl_up = ifelse(increase_cyano_response == "No significant response", "Not significant", 
                                ifelse(increase_cyano_response == "Decrease in cyanobacteria", "Decrease ", 
                                ifelse(increase_cyano_response == "Increase in cyanobacteria", "Increase ", 
                               ifelse(increase_cyano_response == "shift", "Shift", NA))))) %>% 
    mutate(cyano_response_wl_down = ifelse(decrease_cyano_response == "No significant response", "Not significant", 
                                ifelse(decrease_cyano_response == "Decrease in cyanobacteria", "Decrease", 
                                ifelse(decrease_cyano_response == "Increase in cyanobacteria", "Increase", 
                               ifelse(decrease_cyano_response == "shift", "Shift to \n Cyanobacteria \n dominated\n community", NA))))) %>% 
  mutate(combined_cyano_response_binary = case_when(
  is.na(cyano_binary_decrease) & prop_delta_wl_prioritydecrease < 0 ~ NA_real_,
  is.na(cyano_binary_decrease) ~ cyano_binary_increase,
  TRUE ~ cyano_binary_decrease
)) |>
       mutate(combined_cyano_response_binary_increasepriority = case_when(
  prop_delta_wl_priorityincrease < 0 ~ cyano_binary_decrease,
  prop_delta_wl_priorityincrease > 0 ~ cyano_binary_increase,
  TRUE ~ cyano_binary_increase
)) |> 
   mutate(combined_cyano_response = ifelse(is.na(decrease_cyano_response), cyano_response_wl_up, decrease_cyano_response)) %>% 
  mutate(combined_cyano_response_increasepriority = ifelse(is.na(cyano_response_wl_up), decrease_cyano_response, cyano_response_wl_up)) %>% 
    mutate(phyto_response_wl_up = ifelse(increase_phyto_response == "No significant response", "Not significant", 
                                ifelse(increase_phyto_response == "Decrease in phytoplankton", "Decrease ", 
                                ifelse(increase_phyto_response == "Increase in phytoplankton", "Increase ", 
                               ifelse(increase_phyto_response == "shift", "Shift", NA))))) %>% 
    mutate(phyto_response_wl_down = ifelse(decrease_phyto_response == "No significant response", "Not significant", 
                                ifelse(decrease_phyto_response == "Decrease in phytoplankton", "Decrease", 
                                ifelse(decrease_phyto_response == "Increase in phytoplankton", "Increase", 
                               ifelse(decrease_phyto_response == "shift", "Shift to\n Cyanobacteria\n dominated\n community", NA))))) %>% 
  mutate(combined_phyto_response = ifelse(is.na(phyto_response_wl_down), phyto_response_wl_up, phyto_response_wl_down)) |> 
  mutate(combined_phyto_response_increasepriority = ifelse(is.na(phyto_response_wl_up), phyto_response_wl_down, phyto_response_wl_up))
```

## Set up logistic regression for effect of water level change on phytoplankton and cyanobacteria communities

We ran the logistic regression three ways. In the manuscript, the logistic regression includes both water level increases and decreases. We also prioritized only water level increases (any lake that had both a water level increase and decrease was kept as only an increase) and subsequently prioritized only water level decreases (any lake that had both a water level increase and decrease was kept as only a decrease)

```{r}
phytoincrease_logreg <- boxplotdata2_opposite |> 
  dplyr::select(prop_1delta_wl_decrease, prop_1delta_wl_increase, phyto_binary_increase, phyto_binary_decrease, phyto_response_wl_up, phyto_response_wl_down, name_yr, trophic_status_reduced) |> 
  #mutate(phyto_binary_decrease = NA) |> 
  filter(!is.na(phyto_response_wl_up)) |> 
  filter(!is.na(prop_1delta_wl_increase))

phytodecrease_logreg <- boxplotdata2_opposite |> 
  dplyr::select(prop_1delta_wl_decrease, prop_1delta_wl_increase, phyto_binary_increase, phyto_binary_decrease, phyto_response_wl_up, phyto_response_wl_down, name_yr, trophic_status_reduced) |> 
  filter(!is.na(phyto_response_wl_down)) |> 
  filter(!is.na(prop_1delta_wl_decrease))

phyto_logreg <- rbind(phytodecrease_logreg, phytoincrease_logreg)

phytos <- phyto_logreg |> 
  mutate(response = as.numeric(ifelse(!is.na(prop_1delta_wl_decrease), phyto_binary_decrease, 
                                 ifelse(!is.na(prop_1delta_wl_increase), phyto_binary_increase, NA))), 
         deltawl = as.numeric(ifelse(!is.na(prop_1delta_wl_decrease), prop_1delta_wl_decrease, 
                                 ifelse(!is.na(prop_1delta_wl_increase), prop_1delta_wl_increase, NA)))) |> 
  dplyr::select(response, deltawl, name_yr, trophic_status_reduced) |> 
  mutate(type = "phyto")


ggplot(phytos, aes(x = deltawl, y = response)) +
  geom_jitter(height = 0.05, width = 0, alpha = 0.4) +
  stat_smooth(method = "glm",
              method.args = list(family = "binomial"),
              se = TRUE) + labs(
    x = "Proportional Water Level Change",
    y = "Probability of Phytoplankton Increase") +
  theme_classic()

cyanoincrease_logreg <- boxplotdata2_opposite |> 
  dplyr::select(prop_1delta_wl_decrease, prop_1delta_wl_increase, cyano_binary_increase, cyano_binary_decrease, cyano_response_wl_up, cyano_response_wl_down, name_yr, trophic_status_reduced) |> 
  mutate(cyano_binary_decrease = NA) |> 
  filter(!is.na(cyano_response_wl_up)) |> 
  filter(!is.na(prop_1delta_wl_increase))

cyanodecrease_logreg <- boxplotdata2_opposite |> 
  dplyr::select(prop_1delta_wl_decrease, prop_1delta_wl_increase, cyano_binary_increase, cyano_binary_decrease, cyano_response_wl_up, cyano_response_wl_down, name_yr, trophic_status_reduced) |> 
  filter(!is.na(cyano_response_wl_down)) |> 
  filter(!is.na(prop_1delta_wl_decrease))

cyano_logreg <- rbind(cyanodecrease_logreg, cyanoincrease_logreg)

cyanos <- cyano_logreg |> 
  mutate(
    response = as.numeric(
      ifelse(!is.na(prop_1delta_wl_decrease), cyano_binary_decrease, 
      ifelse(!is.na(prop_1delta_wl_increase), cyano_binary_increase, NA))
    ), 
    deltawl = as.numeric(
      ifelse(!is.na(prop_1delta_wl_decrease), prop_1delta_wl_decrease, 
      ifelse(!is.na(prop_1delta_wl_increase), prop_1delta_wl_increase, NA))
    )
  ) |>
  mutate(type = "cyano") |> 
  dplyr::select(response, deltawl, name_yr, trophic_status_reduced, type)


ggplot(cyanos, aes(x = deltawl, y = response)) +
  geom_jitter(height = 0.05, width = 0, alpha = 0.4) +
  stat_smooth(method = "glm",
              method.args = list(family = "binomial"),
              se = TRUE) +
  labs(
    x = "Proportional Water Level Change",
    y = "Probability of Cyanobacteria Increase"
  ) +
  theme_classic()

all_logreg <- rbind(phytos, cyanos)


ggplot(all_logreg, aes(x = deltawl, y = response)) +
  geom_point(aes(color = trophic_status_reduced), alpha = 0.6, size = 4) +
  geom_smooth(
    method = "glm",
    method.args = list(family = "binomial"),
    se = FALSE,
    color = "black"
  ) +
  facet_wrap(~type) +
  theme_bw() +
  labs(
    x = "Proportional Water Level Change",
    y = "Probability of Increase"
  )

logregform_phyto <-  glm(formula=response ~ deltawl, data=phytos, family=binomial)
summary(logregform_phyto)

logregform_cyano <-  glm(formula=response ~ deltawl, data=cyanos, family=binomial)
summary(logregform_cyano)

```

## Logistic Regression,Non Significant responses excluded

```{r}

boxplotdata2_opposite <- boxplotdata2 %>% 
  #mutate(prop_delta_wl_prioritydecrease, = delta_wl_direction/depth_max_m) %>% 
    mutate(phyto_binary_increase = ifelse(increase_phyto_response == "No significant response", NA, 
                                ifelse(increase_phyto_response == "Decrease in phytoplankton", 0, 
                                ifelse(increase_phyto_response == "Increase in phytoplankton", 1, NA)))) %>% 
  mutate(phyto_binary_decrease = ifelse(decrease_phyto_response == "No significant response", NA, 
                                ifelse(decrease_phyto_response == "Decrease in phytoplankton", 0, 
                                ifelse(decrease_phyto_response == "Increase in phytoplankton", 1, NA)))) %>%
  # mutate(combined_phyto_response_text = decrease_phyto_response %>% 
  #          is.na() %>% 
  #          ifelse(increase_phyto_response, decrease_phyto_response)) %>% 
    mutate(combined_phyto_response_binary = case_when(
  is.na(phyto_binary_decrease) & prop_delta_wl_prioritydecrease < 0 ~ NA_real_,
  is.na(phyto_binary_decrease) ~ phyto_binary_increase,
  TRUE ~ phyto_binary_decrease
)) |> 
    mutate(combined_phyto_response_binary_increasepriority = case_when(
 prop_delta_wl_priorityincrease < 0 ~ phyto_binary_decrease,
  is.na(phyto_binary_increase) & prop_delta_wl_priorityincrease > 0 ~ phyto_binary_increase,
  TRUE ~ phyto_binary_increase
)) |> 
    #mutate(combined_phyto_response_binary = ifelse(is.na(phyto_binary_decrease), phyto_binary_increase, phyto_binary_decrease)) %>% 
    mutate(cyano_binary_increase = ifelse(increase_cyano_response == "No significant response", NA, 
                                ifelse(increase_cyano_response == "Decrease in cyanobacteria", 0, 
                                ifelse(increase_cyano_response == "Increase in cyanobacteria", 1, NA)))) %>% 
  mutate(cyano_binary_decrease = ifelse(decrease_cyano_response == "No significant response", NA, 
                                ifelse(decrease_cyano_response == "Decrease in cyanobacteria", 0, 
                                ifelse(decrease_cyano_response == "Increase in cyanobacteria", 1, NA)))) %>%
  # mutate(combined_phyto_response_text = decrease_phyto_response %>% 
  #          is.na() %>% 
  #          ifelse(increase_phyto_response, decrease_phyto_response)) %>% 
    #mutate(combined_cyano_response_binary = ifelse(is.na(cyano_binary_decrease), cyano_binary_increase, cyano_binary_decrease)) %>% 
  mutate(cyano_response_wl_up = ifelse(increase_cyano_response == "No significant response", "Not significant", 
                                ifelse(increase_cyano_response == "Decrease in cyanobacteria", "Decrease ", 
                                ifelse(increase_cyano_response == "Increase in cyanobacteria", "Increase ", 
                               ifelse(increase_cyano_response == "shift", "Shift", NA))))) %>% 
    mutate(cyano_response_wl_down = ifelse(decrease_cyano_response == "No significant response", "Not significant", 
                                ifelse(decrease_cyano_response == "Decrease in cyanobacteria", "Decrease", 
                                ifelse(decrease_cyano_response == "Increase in cyanobacteria", "Increase", 
                               ifelse(decrease_cyano_response == "shift", "Shift to \n Cyanobacteria \n dominated\n community", NA))))) %>% 
  mutate(combined_cyano_response_binary = case_when(
  is.na(cyano_binary_decrease) & prop_delta_wl_prioritydecrease < 0 ~ NA_real_,
  is.na(cyano_binary_decrease) ~ cyano_binary_increase,
  TRUE ~ cyano_binary_decrease
)) |>
       mutate(combined_cyano_response_binary_increasepriority = case_when(
  prop_delta_wl_priorityincrease < 0 ~ cyano_binary_decrease,
  prop_delta_wl_priorityincrease > 0 ~ cyano_binary_increase,
  TRUE ~ cyano_binary_increase
)) |> 
   mutate(combined_cyano_response = ifelse(is.na(decrease_cyano_response), cyano_response_wl_up, decrease_cyano_response)) %>% 
  mutate(combined_cyano_response_increasepriority = ifelse(is.na(cyano_response_wl_up), decrease_cyano_response, cyano_response_wl_up)) %>% 
    mutate(phyto_response_wl_up = ifelse(increase_phyto_response == "No significant response", "Not significant", 
                                ifelse(increase_phyto_response == "Decrease in phytoplankton", "Decrease ", 
                                ifelse(increase_phyto_response == "Increase in phytoplankton", "Increase ", 
                               ifelse(increase_phyto_response == "shift", "Shift", NA))))) %>% 
    mutate(phyto_response_wl_down = ifelse(decrease_phyto_response == "No significant response", "Not significant", 
                                ifelse(decrease_phyto_response == "Decrease in phytoplankton", "Decrease", 
                                ifelse(decrease_phyto_response == "Increase in phytoplankton", "Increase", 
                               ifelse(decrease_phyto_response == "shift", "Shift to\n Cyanobacteria\n dominated\n community", NA))))) %>% 
  mutate(combined_phyto_response = ifelse(is.na(phyto_response_wl_down), phyto_response_wl_up, phyto_response_wl_down)) |> 
  mutate(combined_phyto_response_increasepriority = ifelse(is.na(phyto_response_wl_up), phyto_response_wl_down, phyto_response_wl_up))

cyanodecreasedata <- boxplotdata2_opposite %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_priorityincrease, combined_cyano_response_increasepriority, combined_cyano_response_binary_increasepriority, name_yr) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>%
  filter(!is.na(combined_cyano_response_binary_increasepriority)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_binaryincrease = combined_cyano_response_binary_increasepriority, 
         response_increase = combined_cyano_response_increasepriority) %>% 
  mutate(type = "cyano")

cyanoincreasedata <- boxplotdata2_opposite %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_priorityincrease, combined_cyano_response_increasepriority, combined_cyano_response_binary_increasepriority, name_yr) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>%
  filter(!is.na(combined_cyano_response_binary_increasepriority)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_binaryincrease = combined_cyano_response_binary_increasepriority, 
         response_increase = combined_cyano_response_increasepriority) %>% 
  mutate(type = "cyano")

phytoincreasedata <- boxplotdata2_opposite %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_priorityincrease, combined_phyto_response_increasepriority, combined_phyto_response_binary_increasepriority, name_yr) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>% 
  filter(!is.na(combined_phyto_response_binary_increasepriority)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_binaryincrease = combined_phyto_response_binary_increasepriority, 
         response_increase = combined_phyto_response_increasepriority) %>% 
  mutate(type = "phyto")

phytodecreasedata <- boxplotdata2_opposite %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_priorityincrease, combined_phyto_response_increasepriority, combined_phyto_response_binary_increasepriority, name_yr) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>% 
  filter(!is.na(combined_phyto_response_binary_increasepriority)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_binaryincrease = combined_phyto_response_binary_increasepriority, 
         response_increase = combined_phyto_response_increasepriority) %>% 
  mutate(type = "phyto")
  
cyanodata <- boxplotdata2_opposite %>% 
  select(prop_delta_wl_priorityincrease, combined_cyano_response_increasepriority, combined_cyano_response_binary_increasepriority, name_yr, trophic_status_reduced) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>%
  filter(!is.na(combined_cyano_response_binary_increasepriority)) %>% 
  # filter(prop_delta_wl<0) %>% 
  rename(response_binary = combined_cyano_response_binary_increasepriority, 
         response = combined_cyano_response_increasepriority) %>% 
  mutate(type = "cyano") |> 
  mutate(wl_priority = "increase")

phytodata <- boxplotdata2_opposite %>% 
  select(prop_delta_wl_priorityincrease, combined_phyto_response_increasepriority, combined_phyto_response_binary_increasepriority, name_yr, trophic_status_reduced) %>% 
  filter(!is.na(prop_delta_wl_priorityincrease)) %>% 
  filter(!is.na(combined_phyto_response_binary_increasepriority)) %>% 
  # filter(prop_delta_wl<0) %>% 
  rename(response_binary = combined_phyto_response_binary_increasepriority, 
         response = combined_phyto_response_increasepriority) %>% 
  mutate(type = "phyto")|> 
  mutate(wl_priority = "increase")
# logregdata <- rbind(cyanoincreasedata, cyanodecreasedata, phytoincreasedata, phytodecreasedata)
logregdata_increase <- rbind(cyanodata, phytodata)
#"#7A524B","#3C7F29","#8DBAE9"

cyanodecreasedata <- boxplotdata2_opposite %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_prioritydecrease, combined_cyano_response, combined_cyano_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl_prioritydecrease)) %>%
  filter(!is.na(combined_cyano_response_binary)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_binary = combined_cyano_response_binary, 
         response = combined_cyano_response) %>% 
  mutate(type = "cyano")

cyanoincreasedata <- boxplotdata2_opposite %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_prioritydecrease, combined_cyano_response, combined_cyano_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl_prioritydecrease)) %>%
  filter(!is.na(combined_cyano_response_binary)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_binary = combined_cyano_response_binary, 
         response = combined_cyano_response) %>% 
  mutate(type = "cyano")

phytoincreasedata <- boxplotdata2_opposite %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_prioritydecrease, combined_phyto_response, combined_phyto_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl_prioritydecrease)) %>% 
  filter(!is.na(combined_phyto_response_binary)) %>% 
  filter(prop_alldelta_wl_increase>0) %>% 
  rename(response_binary = combined_phyto_response_binary, 
         response = combined_phyto_response) %>% 
  mutate(type = "phyto")

phytodecreasedata <- boxplotdata2_opposite %>% 
  select(prop_alldelta_wl_decrease, prop_alldelta_wl_increase, prop_delta_wl_prioritydecrease, combined_phyto_response, combined_phyto_response_binary, name_yr) %>% 
  filter(!is.na(prop_delta_wl_prioritydecrease)) %>% 
  filter(!is.na(combined_phyto_response_binary)) %>% 
  filter(prop_alldelta_wl_decrease<0) %>% 
  rename(response_binary = combined_phyto_response_binary, 
         response = combined_phyto_response) %>% 
  mutate(type = "phyto")
  
cyanodata <- boxplotdata2_opposite %>% 
  select(prop_delta_wl_prioritydecrease, combined_cyano_response, combined_cyano_response_binary, name_yr, trophic_status_reduced) %>% 
  filter(!is.na(prop_delta_wl_prioritydecrease)) %>%
  filter(!is.na(combined_cyano_response_binary)) %>% 
  # filter(prop_delta_wl_prioritydecrease<0) %>% 
  rename(response_binary = combined_cyano_response_binary, 
         response = combined_cyano_response) %>% 
  mutate(type = "cyano") |> 
  mutate(wl_priority = "decrease")

phytodata <- boxplotdata2_opposite %>% 
  select(prop_delta_wl_prioritydecrease, combined_phyto_response, combined_phyto_response_binary, name_yr, trophic_status_reduced) %>% 
  filter(!is.na(prop_delta_wl_prioritydecrease)) %>% 
  filter(!is.na(combined_phyto_response_binary)) %>% 
  # filter(prop_delta_wl_prioritydecrease,<0) %>% 
  rename(response_binary = combined_phyto_response_binary, 
         response = combined_phyto_response) %>% 
  mutate(type = "phyto") |> 
  mutate(wl_priority = "decrease")

# logregdata <- rbind(cyanoincreasedata, cyanodecreasedata, phytoincreasedata, phytodecreasedata)
logregdata_decrease <- rbind(cyanodata, phytodata)
logregdata <- logregdata_decrease
#"#7A524B","#3C7F29","#8DBAE9"

logregdata_increase1 <- logregdata_increase |>
  rename(prop_delta_wl = prop_delta_wl_priorityincrease)

new_all_noNA <- bind_rows(logregdata_decrease, logregdata_increase1) |>
  select(-c(wl_priority)) |>
  unique()

all <- new_all_noNA %>% ggplot(aes(x = prop_delta_wl, y = response_binary, color = as.factor(type), shape = as.factor(type))) +
  scale_color_manual(values = c("#88CCEE", "#117733"), name = "")+
  scale_shape_manual(values = c(8, 20)) +
  geom_point(size = 3) +
  geom_smooth(method = "glm",
              method.args = list(family = "binomial"),
              se = FALSE, linewidth = 3) +
  #geom_jitter()+
 ylab("Response to changes in water level") +
  xlab("Proportion water level change") +
  ggtitle("Combined increases and decreases") + 
  theme_bw() +
  theme(axis.text.x = element_text(hjust = 1, size = 10),
        axis.title.y=element_text(angle=90, vjust = .5, size = 12),
        legend.title=element_blank())
all
#117733
#factor(trophic_status_reduced, c("oligotrophic", "mesotrophic", "eutrophic", "hypereutrophic"))

logregdata_phyto <- new_all_noNA %>% subset(type == "phyto")
logregdata_cyano <- new_all_noNA %>% subset(type == "cyano")

logregform_phyto <-  glm(formula=response_binary ~ prop_delta_wl, data=logregdata_phyto, family=binomial)
summary(logregform_phyto)

logregform_cyano <-  glm(formula=response_binary ~ prop_delta_wl, data=logregdata_cyano, family=binomial)
summary(logregform_cyano)

ggsave(paste0(here(),"/Supplement_figures/logreg_ns.jpg"), plot = all, height = 4, width = 8, units = "in")

```
